---
title: "DATA 621- HW4"
author: "Jose Mawyin"
date: "3/27/2020"
output: pdf_document
---
```{r}
install.packages("Amelia")
```


```{r}
library(ggplot2)
library(corrr)
library(caret) #Marchine and glm library
library(magrittr)
library(dplyr)
library(car)
library(Amelia) #missmap
```


##DATA 621- HW4


Getting data


Overview:


In this homework assignment, you will explore, analyze and model a data set containing approximately 8000 records representing a customer at an auto insurance company. Each record has two response variables. The first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero if the person did not crash their car. But if they did crash their car, this number will be a value greater than zero.
Your objective is to build multiple linear regression and binary logistic regression models on the training data to predict the probability that a person will crash their car and also the amount of money it will cost if the person does crash their car. You can only use the variables given to you (or variables that you derive from the variables provided). Below is a short description of the variables of interest in the data set:


Deliverables:
* A write-up submitted in PDF format. Your write-up should have four sections. Each one is described below. You may assume you are addressing me as a fellow data scientist, so do not need to shy away from technical details.
* Assigned predictions (probabilities, classifications, cost) for the evaluation data set. Use 0.5 threshold.
* Include your R statistical programming code in an Appendix.



Write Up:

1. DATA EXPLORATION (25 Points)
Describe the size and the variables in the insurance training data set. Consider that too much detail will cause a manager to lose interest while too little detail will make the manager consider that you aren’t doing your job. Some suggestions are given below. Please do NOT treat this as a check list of things to do to complete the assignment. You should have your own thoughts on what to tell the boss. These are just ideas.

```{r}
t_location <- "/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /621/621HW4/insurance_training_data.csv"
e_location <- "/Users/josemawyin/Library/Mobile Documents/com~apple~CloudDocs/Data Science Masters /621/621HW4/insurance-evaluation-data.csv"


training <- read.csv(t_location, header = TRUE)
evaluation <- read.csv(e_location, header = TRUE)
```

```{r}
head(training)
```

```{r}
head(evaluation)
table(evaluation$TARGET_FLAG)
```



```{r}
dim(training); dim(evaluation)
```

####Functions
```{r}
#Transform from dollar price factor to numeric
DollarToNum <- function(dollar) {
  Num <- as.numeric(gsub('[$,]', '', dollar))
  return(Num)
}
```


```{r}
cat("Transforming Dollar Price Factors to Numerical\n")
training.2 <- training
training.2$INCOME <- DollarToNum(training.2$INCOME)
training.2$HOME_VAL <- DollarToNum(training.2$HOME_VAL)
training.2$BLUEBOOK <- DollarToNum(training.2$BLUEBOOK)
training.2$OLDCLAIM <- DollarToNum(training.2$OLDCLAIM)
training.2$TARGET_FLAG <- as.factor(training.2$TARGET_FLAG)
training.2$TARGET_FLAG <- make.names(training.2$TARGET_FLAG) %>% as.factor()
table(training.2$TARGET_FLAG)
head(training.2)
```

```{r}
cat("Transforming Dollar Price Factors to Numerical\n")
evaluation.2 <- evaluation
evaluation.2$INCOME <- DollarToNum(evaluation.2$INCOME)
evaluation.2$HOME_VAL <- DollarToNum(evaluation.2$HOME_VAL)
evaluation.2$BLUEBOOK <- DollarToNum(evaluation.2$BLUEBOOK)
evaluation.2$OLDCLAIM <- DollarToNum(evaluation.2$OLDCLAIM)
evaluation.2$TARGET_FLAG <- as.factor(evaluation.2$TARGET_FLAG)
head(evaluation.2)
```


```{r}
cat("How many cases of accepted or denied are thre in the training dataset.\n")
table(training.2$TARGET_FLAG)
```


a. Mean / Standard Deviation / Median
b. Bar Chart or Box Plot of the data
c. Is the data correlated to the target variable (or to other variables?)
```{r}
#cor(training, use="pairwise.complete.obs")
x <- training.2 %>% 
  correlate() %>% 
  focus(training$TARGET_AMT)
```

d. Are any of the variables missing and need to be imputed “fixed”?

```{r}
cat("Counting na's per column\n")
sapply(training.2, function(x) sum(is.na(x)))
cat("% number of rows that have complete observations (no na's in the whole row)\n")
100*sum(complete.cases(training.2))/nrow(training.2)
```
```{r}
missmap(training.2, col=c("red", "black"), legend=FALSE)
```


2. DATA PREPARATION (25 Points)
Describe how you have transformed the data by changing the original variables or creating new variables. If you did transform the data or create new variables, discuss why you did this. Here are some possible transformations.
a. Fix missing values (maybe with a Mean or Median value)
b. Create flags to suggest if a variable was missing
c. Transform data by putting it into buckets
d. Mathematical transforms such as log or square root (or use Box-Cox)
e. Combine variables (such as ratios or adding or multiplying) to create new variables
*How to impute data*
*https://amunategui.github.io/binary-outcome-modeling/*

```{r}
training.3 <- training.2
training.3$AGE[is.na(training.3$AGE)] <- median(training.3$AGE, na.rm=T)
training.3$YOJ[is.na(training.3$YOJ)] <- median(training.3$YOJ, na.rm=T)
training.3$INCOME[is.na(training.3$INCOME)] <- median(training.3$INCOME, na.rm=T)
training.3$HOME_VAL[is.na(training.3$HOME_VAL)] <- median(training.3$HOME_VAL, na.rm=T)
training.3$CAR_AGE[is.na(training.3$CAR_AGE)] <- median(training.3$CAR_AGE, na.rm=T)

cat("Counting na's per column\n")
sapply(training.3, function(x) sum(is.na(x)))
cat("% number of rows that have complete observations (no na's in the whole row)\n")
100*sum(complete.cases(training.3))/nrow(training.3)
```

```{r}
evaluation.3 <- evaluation.2
evaluation.3$AGE[is.na(evaluation.3$AGE)] <- median(evaluation.3$AGE, na.rm=T)
evaluation.3$YOJ[is.na(evaluation.3$YOJ)] <- median(evaluation.3$YOJ, na.rm=T)
evaluation.3$INCOME[is.na(evaluation.3$INCOME)] <- median(evaluation.3$INCOME, na.rm=T)
evaluation.3$HOME_VAL[is.na(evaluation.3$HOME_VAL)] <- median(evaluation.3$HOME_VAL, na.rm=T)
evaluation.3$CAR_AGE[is.na(evaluation.3$CAR_AGE)] <- median(evaluation.3$CAR_AGE, na.rm=T)

cat("Counting na's per column\n")
sapply(evaluation.3, function(x) sum(is.na(x)))
cat("% number of rows that have complete observations (no na's in the whole row)\n")
100*sum(complete.cases(evaluation.3))/nrow(evaluation.3)
```

```{r}
str(training.3)
```

```{r}
#Dropping the index column
only.factor <- training.3 %>% Filter(f = is.factor)
only.num <- training.3 %>% Filter(f = is.numeric)
str(only.factor)
str(only.num)
ncol(training.3)-(ncol(only.factor)+ncol(only.num))
```

```{r}
training.4 <- only.num

#We only are scaling some of the numerical columns in the data set
training.4[c(4, 6,7,8,9,10,11,12,14,15)] <- scale(training.4[c(4, 6,7,8,9,10,11,12,14,15)])

cols.dont.want <- c("INDEX","TARGET_AMT") # if you want to remove multiple columns
training.4 <- training.4[, ! names(training.4) %in% cols.dont.want, drop = F]

training.4$TARGET_FLAG <- training.3$TARGET_FLAG
str(training.4)
```


3. BUILD MODELS (25 Points)
Using the training data set, build at least two different multiple linear regression models and three different binary logistic regression models, using different variables (or the same variables with different transformations). You may select the variables manually, use an approach such as Forward or Stepwise, use a different approach such as trees, or use a combination of techniques. Describe the techniques you used. If you manually selected a variable for inclusion into the model or exclusion into the model, indicate why this was done.
Discuss the coefficients in the models, do they make sense? For example, if a person has a lot of traffic tickets, you would reasonably expect that person to have more car crashes. If the coefficient is negative (suggesting that the person is a safer driver), then that needs to be discussed. Are you keeping the model even though it is counter intuitive? Why? The boss needs to know.

```{r}
cat("Showing the class type for all the columns in the dataset.\n")
sapply(training.4, class)
```

Splitting Training set into training/testing dataframes:

```{r}
set.seed(1234)
#training.4 <- only.num
splitIndex <- createDataPartition(training.4$TARGET_FLAG, p = .7, list = FALSE, times = 1)
training.4$TARGET_FLAG <- training.3$TARGET_FLAG
train.df_num <- training.4[ splitIndex,]
test.df_num  <- training.4[-splitIndex,]
training.5 <- only.factor
training.5$TARGET_FLAG <- training.3$TARGET_FLAG
train.df_fac <- training.5[splitIndex,]
test.df_fac <- training.5[-splitIndex,]
```

```{r}
dim(train.df_num); dim(train.df_num)
```


######################
Linear Regression Models
######################

### Only Nums LM
```{r}
head(train.df_num); head(test.df_num)
```


```{r}
set.seed(1234)

train.df.lm <- train.df_num
train.df.lm$TARGET_FLAG <-as.numeric(train.df.lm$TARGET_FLAG)
head(train.df.lm)
# Set up a 10-fold cross validation
tc <- trainControl(method = "cv", number = 10)
# Include the setup in your model
lm1_cv <- train(form = TARGET_FLAG ~ ., data = train.df.lm, method = "lm",
             trControl = tc) # here
lm1_cv
```

```{r}
test.df.lm <- test.df_num
test.df.lm$TARGET_FLAG <-as.numeric(test.df.lm$TARGET_FLAG)
predlm1 = predict(lm1_cv, newdata=test.df.lm)
head(predlm1)
confusionMatrix(data=predlm1, test.df.lm$TARGET_FLAG)
```

### Only Factors LM

```{r}
set.seed(1234)
#Power Component
train.df.lm2 <- train.df_fac
train.df.lm2$TARGET_FLAG <-as.numeric(train.df.lm2$TARGET_FLAG)
# Set up a 10-fold cross validation
tc <- trainControl(method = "cv", number = 10)
# Include the setup in your model
lm2_cv <- train(form = TARGET_FLAG ~ ., data = train.df.lm2, method = "lm",
             trControl = tc) # here
lm2_cv
```


### All Predictors LM

```{r}
set.seed(1234)
#Power Component
train.df.lm3 <- cbind(train.df_num,train.df_fac)
train.df.lm3$TARGET_FLAG <-as.numeric(train.df.lm3$TARGET_FLAG)
# Set up a 10-fold cross validation
tc <- trainControl(method = "cv", number = 10)
# Include the setup in your model
lm3_cv <- train(form = TARGET_FLAG ~ ., data = train.df.lm3, method = "lm",
             trControl = tc) # here
lm3_cv
summary(lm3_cv)
```



```{r}
str(train.df.lm)
table(train.df.lm$CLM_FREQ)
```

```{r}
featurePlot(x = only.factor[, 2:11], 
            y = training.3$TARGET_FLAG, 
            plot = "pairs",
            strip=strip.custom(par.strip.text=list(cex=.7)),
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")))
```

```{r}
featurePlot(x = only.num[, 3:15], 
            y = training.3$TARGET_FLAG, 
            plot = "box",
            strip=strip.custom(par.strip.text=list(cex=.7)),
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")))
```

```{r}
featurePlot(x = train.df[, 1:13], 
            y = train.df$TARGET_FLAG, 
            plot = "density",
            strip=strip.custom(par.strip.text=list(cex=.7)),
            scales = list(x = list(relation="free"), 
                          y = list(relation="free")))
```


#####
Power Transform
#####


```{r}
summary(p1 <- powerTransform(TARGET_FLAG ~ ., train.df.lm))
# fit linear model with transformed response:
coef(p1, round=TRUE)
summary(lm1 <- lm(bcPower(TARGET_FLAG, p1$roundlam) ~ ., train.df.lm))
summary(lm2 <- lm(TARGET_FLAG ~ ., train.df.lm))
```

######################
Logistic Regression Models
######################
```{r}
#mylogit <- glm(TARGET_FLAG ~ AGE + INCOME + HOME_VAL, data = training, family = "binomial")
logit.1 <- glm(TARGET_FLAG ~ AGE + INCOME + HOME_VAL, data = train.df, family = "binomial")
summary(logit.1)
```

```{r}
ctrl <- trainControl(method = "repeatedcv", number = 10, savePredictions = TRUE)

mod_fit <- train(form = TARGET_FLAG ~ ., data = train.df, method="glm", family="binomial", trControl = ctrl, tuneLength = 5, na.action = na.pass)

pred = predict(mod_fit, newdata=test.df)
head(pred)
confusionMatrix(data=pred, test.df$TARGET_FLAG)
```

```{r}
set.seed(12345)
knnctrl <- trainControl(method="repeatedcv",
                        repeats = 3,
                        summaryFunction=twoClassSummary, 
                        classProbs=T,
                        savePredictions = T) #,classProbs=TRUE,summaryFunction = twoClassSummary)
knnFit <- train(form = TARGET_FLAG ~ ., data = train.df, method = "knn", trControl = knnctrl, tuneLength = 10)

#Output of kNN fit
knnFit
plot(knnFit)
```

```{r}
knnPredict <- predict(knnFit,newdata = test.df )
#Get the confusion matrix to see accuracy value and other parameter values
knnCM <- confusionMatrix(knnPredict, test.df$TARGET_FLAG )
knnCM
```



=========
Random Forest
=========

```{r}
set.seed(12311)
treectrl <- trainControl(method="repeatedcv", number=10, repeats=3)
#treectrl <- rpart.control(minsplit =1,minbucket=1, cp=0)
treeFit <-
  train(form = TARGET_FLAG ~ ., data = train.df,
    method = "rpart",
    metric = "Accuracy",
    trControl = treectrl,
    cp=0.002,maxdepth=8,
    parms = list(split = "information")
  )

formula <- as.formula(TARGET_FLAG ~.)
#t <- train(formula,HW2_data.Train,method = "rpart",cp=0.002,maxdepth=8)
```

```{r}
treeFit
```

```{r}
plot(treeFit$finalModel, uniform=TRUE,
     main="Classification Tree")
text(treeFit$finalModel, use.n.=TRUE, all=TRUE, cex=.8)
```

```{r}
treePredict <- predict(treeFit,newdata = test.df )
#Get the confusion matrix to see accuracy value and other parameter values
treeCM <- confusionMatrix(treePredict, test.df$TARGET_FLAG )
treeCM
```

4. SELECT MODELS (25 Points)
Decide on the criteria for selecting the best multiple linear regression model and the best binary logistic regression model. Will you select models with slightly worse performance if it makes more sense or is more parsimonious? Discuss why you selected your models.
For the multiple linear regression model, will you use a metric such as Adjusted R2, RMSE, etc.? Be sure to explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other relevant model output. Using the training data set, evaluate the multiple linear regression model based on (a) mean squared error, (b) R2, (c) F-statistic, and (d) residual plots. For the binary logistic regression model, will you use a metric such as log likelihood, AIC, ROC curve, etc.? Using the training data set, evaluate the binary logistic regression model based on (a) accuracy, (b) classification error rate, (c) precision, (d) sensitivity, (e) specificity, (f) F1 score, (g) AUC, and (h) confusion matrix. Make predictions using the evaluation data set.



Practical guide to implement machine learning with CARET package in R (with practice problem)
https://www.analyticsvidhya.com/blog/2016/12/practical-guide-to-implement-machine-learning-with-caret-package-in-r-with-practice-problem/








